---
title: 'Reproducible Research: Peer Assessment 1'
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

## Preparing the environment

```{r, echo = TRUE, results = 'hide', message=FALSE}
library(dplyr)
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

## Loading and preprocessing the data

```{r}
if(!file.exists("activity.zip")) {
      fileUrl <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(fileUrl, "activity.zip", method = "curl")
} 
filename = "activity.csv"

if(!file.exists(filename)) unzip("activity.zip")

activity <- read.csv(filename, header = TRUE, sep = ',', quote = '\"',stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
d_t <- as.POSIXlt(activity$date)
d_t$hour <- activity$interval %/% 100
d_t$min <- activity$interval %% 100
activity$time <- substr(as.POSIXct(d_t), 12, 16)
activity <- tbl_df(activity)
```

## What is mean total number of steps taken per day?

```{r}
totalByDate <- activity %>% group_by(date) %>% summarize(total_steps=sum(steps))

with(totalByDate, plot(date, total_steps, main = 'Total steps per day', ylab = 'Steps', xlab = 'Date'))
mean_steps <- mean(totalByDate$total_steps,na.rm = TRUE)
median_steps <- median(totalByDate$total_steps,na.rm = TRUE)      
abline(h=mean_steps, col = 'green')
abline(h=median_steps, col = 'blue')
```

The mean of steps per day is **`r as.integer(mean_steps) `**
and the median is **`r as.integer(median_steps) `**.

## What is the average daily activity pattern?

```{r}
dailyPattern <- activity %>% group_by(time) %>% summarize(average_steps=mean(steps,na.rm = TRUE))

with(dailyPattern, plot(as.POSIXct(strptime(time, "%H:%M")),
                        average_steps, type = "l", lwd = 3,
                        main = "Daily activity pattern",
                        xlab = "time"))
```

Max number of steps per interval for an average day is 
**`r as.integer(max(dailyPattern$average_steps))`**.
This maxinum is achieved at interval **`r dailyPattern[which.max(dailyPattern$average_steps),]$time`**.


## Imputing missing values

```{r}

```


## Are there differences in activity patterns between weekdays and weekends?

```{r}

```

